---
title: "RealDirect Data Strategy"
author: "Thijs Leufkens"
date: "4-1-2017"
output: html_document
---

First we load the .xls file.
the gdata package is required for this.
Our Excel has some header text on the top, so we add pattern = "BOROUGH" to start at the first relevant row.

```{r loadmanhattanfile, echo=TRUE}
require(gdata)
MH1 <- read.xls("rollingsales_manhattan.xls", pattern = "BOROUGH")
```

a Quick summary of the data:

```{r head}
head(MH1)
summary(MH1)
str(MH1)
```

test
```{r test}
#df.list <- ('a')
#final.df <- ('b')
#str(df.list)
#str(final.df)
#head(final.df)

#substr(final.df$AREA,14, ".xls")

final.df$AREA <- substr(final.df$AREA, 14, (nchar(final.df$AREA)-4))

```

Load multiple files and add filename as area (to be cleaned later):
```{r multiplefiles}
library(plyr)
library(dplyr)
filenames <- list.files(pattern = ".xls")
df.list <- lapply(filenames, function(x) 
  {
    tmp       <- read.xls(x, pattern = "BOROUGH")
    tmp$AREA  <- x
    tmp
  }
  )
final.df <- do.call(rbind, df.list)

```

a Quick summary of the data:

```{r headall}
head(final.df)
summary(final.df)
str(final.df)
```

Most data is stored in factors, let's convert these in more meaningfull types:

```{r cleaning}
final.df$SALE.PRICE <- as.numeric(final.df$SALE.PRICE)
final.df$SALE.DATE <- as.Date(final.df$SALE.DATE)
final.df$LAND.SQUARE.FEET <- as.numeric(final.df$LAND.SQUARE.FEET)
final.df$GROSS.SQUARE.FEET <- as.numeric(final.df$GROSS.SQUARE.FEET)
final.df$RESIDENTIAL.UNITS <- as.integer(final.df$RESIDENTIAL.UNITS)
final.df$TOTAL.UNITS <- as.integer(final.df$TOTAL.UNITS)
final.df$AREA <- substr(final.df$AREA, 14, (nchar(final.df$AREA)-4)) # Remove the first 13 characters and the last 4
```

Some EDA:

```{r EDA}
hist(MH1$SALE.PRICE)
hist(MH1$SALE.PRICE[SALE.PRICE>1])
hist(MH1$GROSS.SQUARE.FEET[GROSS.SQUARE.FEET < 10])
```

looks like we have some missing prices and Square feet numbers
Let's remove the lines with no sales price
```{r realsales}
MH.SALE <- MH1[MH1$SALE.PRICE>1,]
hist(MH.SALE$SALE.PRICE)
plot(MH.SALE$GROSS.SQUARE.FEET,MH.SALE$SALE.PRICE)
plot(log(MH.SALE$GROSS.SQUARE.FEET),log(MH.SALE$SALE.PRICE))
```

```{r check}
summary(MH.SALE$GROSS.SQUARE.FEET)
summary(MH.SALE$SALE.PRICE)
summary(MH.SALE$BUILDING.CLASS.CATEGORY)
```

I checked Elevator appartments first but these seem to have the most missing square feet records, so I switched to family homes:

```{r family}
library(plyr)
MH.FAM <- MH.SALE[which(grepl("FAMILY", MH.SALE$BUILDING.CLASS.CATEGORY)),]

summary(MH.FAM)
hist(log(MH.FAM$SALE.PRICE))
plot(MH.FAM$GROSS.SQUARE.FEET, MH.FAM$SALE.PRICE)
plot(log(MH.FAM$GROSS.SQUARE.FEET),log(MH.FAM$SALE.PRICE))
```

remove values that are outliers according to the book:

```{r outliers}
MH.FAM$outliers <- (log(MH.FAM$SALE.PRICE) <= 5) + 0
MH.FAM <- MH.FAM[which(MH.FAM$outliers==0),]
plot(log(MH.FAM$GROSS.SQUARE.FEET),log(MH.FAM$SALE.PRICE))
```

Linear Regression

```{r linearfamily}
MH.FAM <- MH.FAM[which(MH.FAM$GROSS.SQUARE.FEET>0 & MH.FAM$LAND.SQUARE.FEET>0),]

linfam1 <- lm(log(MH.FAM$SALE.PRICE) ~ log(MH.FAM$GROSS.SQUARE.FEET))
summary(linfam1)

plot(log(MH.FAM$GROSS.SQUARE.FEET),log(MH.FAM$SALE.PRICE))
abline(linfam1,col="red",lwd=2)
plot(resid(linfam1))
```

```{r model2}
model2 <- lm(log(MH.FAM$SALE.PRICE) ~ log(MH.FAM$GROSS.SQUARE.FEET) + log(MH.FAM$LAND.SQUARE.FEET) + factor(MH.FAM$NEIGHBORHOOD),data=MH.FAM)
summary(model2)
plot(resid(model2))
```